---
title: "Lab X - Accessible Data Viz"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(countrycode)
library(ggpattern)
library(ggplot2)
library(RColorBrewer)
```

## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.

### Exercise 2

I.

II.

III.

IV.

### Exercise 3

Below is the code for the initial data visualization:

```{r plot-penguins}
penguins <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')
penguins %>%
  group_by(species) %>%
  summarise(mean_bill = mean(bill_length_mm, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = fct_reorder(species, mean_bill), y = mean_bill, fill = species)) +
  geom_col() + 
  coord_flip() +
  labs(title = "Average Bill Length of Three Penguin Species", x = "Penguin Species", y = "Average Bill Length in Millimeters", fill = "Penguin Species") 
```

Add code and narrative as needed to improve the accessibility of the above data visualization by focusing on the color palette and alternative text.

```{r}

```


### Exercise 4

Fill in the blanks in the code chunk below appropriately:

```{r pattern-penguins}
penguins %>%
  group_by(species) %>%
  summarise(mean_bill = mean(bill_length_mm, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = fct_reorder(species, mean_bill), y = mean_bill, fill = species)) +
  geom_col_pattern(
    aes(pattern = _____ , fill = _____, pattern_fill = _____,
    colour                   = 'black', 
    pattern_density          = 0.35, 
    pattern_key_scale_factor = 1.3) +
  theme_bw() +
  scale_pattern_fill______ + 
  theme(legend.position = 'none') + 
  coord_flip() +
  labs(title = "Average Bill Length of Three Penguin Species", x = "Penguin Species", y = "Average Bill Length in Millimeters", fill = "Penguin Species") 
```

### Exercise 5

Add code and narrative as needed for your group's final data visualization.

```{r}

```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.
