---
title: "Lab - Accessible Data Viz"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

## Load packages and data

```{r load-packages, message=FALSE}

# Use install.packages("package") to install any of the packages below
library(tidyverse)
library(openintro)
library(countrycode)
library(ggpattern)
library(ggplot2)
library(RColorBrewer)
```

```{r load-data}

chocolate <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')


penguins <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')
```


```{r preprocess-data, fig.alt = "Update me with real alt text"}

chocolate <- chocolate %>%
  mutate(continent = countrycode(country_of_bean_origin, 
                                 origin = "country.name", 
                                 destination = "continent")
         )

chocolate %>%
  filter(company_location == "U.S.A.", 
         country_of_bean_origin != "Blend") %>%
  group_by(country_of_bean_origin, continent) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = fct_reorder(country_of_bean_origin,
                                       count), 
                       y = count, 
                       fill = continent)
         ) +
  geom_col() + 
  coord_flip() +
  labs(
    title = "Bean Origin of Chocolate Bars 
            Manufactured in the U.S.A.", 
    y = "Number of Chocolate Bars 
        Manufactured in the U.S.A.", 
    x = "Country of Bean Origin", 
    fill = "Continent") +
  scale_fill_viridis_d() 

```



## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.

### Exercise 2

I. Why did this alternative text meet or not meet the criteria for alternative text?

II. Why did this alternative text meet or not meet the criteria for alternative text?

III. Why did this alternative text meet or not meet the criteria for alternative text?

IV. Why did this alternative text meet or not meet the criteria for alternative text?

### Exercise 3

Below is the code for the initial data visualization:

```{r plot-penguins}

penguins %>%
  group_by(species) %>%
  summarise(mean_bill = mean(bill_length_mm, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = fct_reorder(species, mean_bill), y = mean_bill, fill = species)) +
  geom_col() + 
  coord_flip() +
  labs(title = "Average Bill Length of Three Penguin Species", x = "Penguin Species", y = "Average Bill Length in Millimeters", fill = "Penguin Species") 
```

Add code and narrative as needed to improve the accessibility of the above data visualization by focusing on the color palette and alternative text.

```{r improve-alt-text-color}

```


### Exercise 4

Fill in the blanks in the code chunk below appropriately. Change to eval = TRUE when you are happy with your changes:

```{r pattern-penguins, eval = FALSE}
penguins %>%
  group_by(species) %>%
  summarise(mean_bill = mean(bill_length_mm, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = fct_reorder(species, mean_bill), y = mean_bill, fill = species)) +
  geom_col_pattern(
    aes(pattern = _____ , fill = _____, pattern_fill = _____,
    colour                   = 'black', 
    pattern_density          = 0.35, 
    pattern_key_scale_factor = 1.3) +
  theme_bw() +
  scale_pattern_fill______ + 
  theme(legend.position = 'none') + 
  coord_flip() +
  labs(title = "Average Bill Length of Three Penguin Species", x = "Penguin Species", y = "Average Bill Length in Millimeters", fill = "Penguin Species") 
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.
